On the use of decision trees to identify the local extrema in parallel global optimization algorithms

### 2023-01-31 2023/PAVT/ParallelDecisionTree.tex
add some references

Since in real global optimization problems each computing of the function value (hereafter called  \textit{a trial}) is a computation-costly operation, one has to reduce the number of such trials. It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, in which the solution of the  problem can be found. The global search algorithm (GSA) is based on this idea. In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves pattern search method) to reduce the number of trials executed. The decision on the run of the local method will be made using a decision tree.

The local optimization algorithms are intended for determining only one of local extrema within a set  of feasible solutions, in which the objective function takes the extremal (maximum or minimum) value.  Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished. Note that the application of the first-order methods (not to mention the second one) in the problems  with the ``black-box'' type functions is difficult since the problem of numerical estimating the gradient arises here. Therefore, further we will consider only the zero-order methods.

>>>>

Since in real global optimization problems each computing of the function value (hereafter called  \textit{a trial}) is a computation-costly operation, one has to reduce the number of such trials. It  can be achieved by intentional choice of variants in the course of search for the optimal solution cutting  off unpromising search subdomains and investigating only these ones, in which the solution of the  problem can be found. The global search algorithm (GSA) is based on this idea \cite{Strongin2000}. In  the present work, we tried to combine GSA and a local optimization method (Hooke-Jeeves pattern search method \cite{HookJeeves}) to reduce the number of trials executed. The decision on the run of the local method will be made using a decision tree.

The local optimization algorithms are intended for determining only one of local extrema within a set  of feasible solutions, in which the objective function takes the extremal (maximum or minimum) value.  Among the local methods,  the zero-order methods and the gradient methods (of the 1\textsuperscript{st} order and of the  2\textsuperscript{nd} one) are traditionally distinguished \cite{Nocedal}. Note that the application of the first-order methods (not to mention the second one) in the problems  with the ``black-box'' type functions is difficult since the problem of numerical estimating the gradient arises here \cite{Kelley}. Therefore, further we will consider only the zero-order methods.
